Eclipse Che for Fair and Easy Education

### 2019-05-04 paper.tex
raw transcription of introduction

\begin{abstract}
Die \LaTeX-Klasse \texttt{lni} setzt die Layout-Vorgaben für Beiträge in LNI Konferenzbänden um.
Dieses Dokument beschreibt ihre Verwendung und ist ein Beispiel für die entsprechende Darstellung.
Der Abstract ist ein kurzer Überblick über die Arbeit der zwischen 70 und 150 Wörtern lang sein und das Wichtigste enthalten sollte.
Die Formatierung erfolgt automatisch innerhalb des abstract-Bereichs.
\end{abstract}

\begin{keywords}
LNI Guidelines \and \LaTeX Vorlage
\end{keywords}

\section{Verwendung}
Die GI gibt unter \url{http://www.gi-ev.de/LNI} Vorgaben für die Formatierung von Dokumenten in der LNI Reihe.
Für \LaTeX-Dokumente werden diese durch die Dokumentenklasse \texttt{lni} realisiert.

Dieses Dokument basiert auf der offiziellen Dokumentation, simplifiziert und setzt grundlegendes LaTeX-Wissen voraus.
Es werden generische Platzhalter an die entsprechenden Stellen (wie beispielsweise die Authoren-Angaben) gesetzt und nicht weiter an anderer Stelle dokumentiert.

Dieses Template ist wie folgt gegliedert:
\Cref{sec:demos} zeigt Demonstrationen der LNI-Verlage.
\Cref{sec:lniconformance} zeigt die Einhaltung der Richtlinien durch einfachen Text.

\section{Demonstrationen}
\label{sec:demos}
Das Symbol für Potenzmengen ($\powerset$) wird korrekt angezeigt.
Es ist kein Weierstraß-p ($\wp$) mehr.

>>>>

\begin{abstract}
Eclipse Che is a revolutionary cloud-based IDE, inheriting the approach of building cloud-ready software within a 
remote environment. 
Instead of developing on your local machine, a remote development-server is accessed via a webbrowser 
and the artifacts are compiled into docker-containers.

This paper shortly summarizes the required environment, introduces the concept and workflow of Eclipse Che and evaluates 
the benefits and downsides of this approach. 
Focus is set on a new point of view, enhancing computer studies and 
programming classes with this technology. 
\end{abstract}

\begin{keywords}
Eclipse Che, Cloud-Ready, Cloud-Native, DevOps, Docker, Kubernetes, OpenShift
\end{keywords}

\section{Introduction}
As children we, used to look into the clouds and count sheep. 
As developers, we look into the clouds and check our applications.  
Cloud technologies are already state of the art, enabling global players like Google to run applications on a world-scale.
The market-leaders got their portfolio stocked up with  a  variety  of  different  cloud  technologies,  ranging 
from cloud computing, cloud storage, database-as-a-service or rentable authentication. 
Modern Software needs to be scalable, self-contained, and should run on any device, aswell as in the cloud.  
This trend is also visible in Googles searches (TODO: Abbildung 1).

Figure 1(TODO LINK) shows the steadily rising interest in the most popular cloud technologies. 
Docker is a container-engine enabling OS-Virtualization, Kubernetes is a Container-Platform manager and Openshift is a software-suite based on Kubernetes. 
These technologies will be summarized in 2.1.  
As cloud-technology such technologies are referred which achieve five attributes: 
On-demand self-service, broad network-access, resource pooling, rapid elasticity and measured service [15]. 
Cloud technology usually splits into two components:  Infrastructure  and  Services,  where  docker and Kubernetes are considered infrastructure and provide a  platform  to  host  services.  
Most  of  the  attributes  are (pseudo-) achievable by the infrastructure alone – such as restarting containers. 
Software running as a service can be put into two categories: 
“Cloud-Ready” – meaning that it can be run in common cloud environments, such as Kubernetes. 
The term cloudready  does  usually  not  include  live-scalability. Cloud-readieness is mostly achieved by removing dependencies and either running on bare-linux or being ported into a dockercontainer. 
Lately the term “Cloud-native” came up [17], referring to software which is built from the very first steps to run inside the cloud. This kind of software is usually running inside a docker-container and both resilient and elasticby design [8].  
Resilience refers to the ability of handling failure, both from external sources and restarting/configuring themselves.  
Elasticy is the ability to request more resources on demand, performing more computations when required, but also to free the resources if not needed. 
Cloud-ready Software usually achieves only a certain degree of resilience.   
Cloud-Native applications are aware of their context and collect meaningful metrics for the platform that they cannot reach other services or need more capacities [16]. 
The applications must also be able to use the gained resources in a productive and healthy way.  
Modern  cloud-platforms  can  perform  pseudo-scalabilty for cloud-ready-services by starting multiple instances of the same service. 
The cloud-platform performs load-balancing between these instances. 
This naive approach can work out fine, but for example starting two services working on the same database will not yield any real scalability. 
Another common problem arises when two services are started on the same machine and need the same port, just to name two problems with cloud-ready  software in a cloud environment.

As Cloud is a central element of modern IT-development, it is strange to see that it is not yet widely teached across 
universities in Germany \footnote{https://www.hochschulkompass.de/studium/studiengangsuche lists currently only 10 german universities with courses on cloud computing, about 30 courses on virtualization }.
There are courses to teach either classic virtualization or docker-basics, but the knowledge of full-stack cloudplatforms and cloud-native development is sealed behind corporate doors and conference workshops.
Meanwhile,  the  annual  Stackoverflow  developersurvey [22] showed, that DevOps-engineers are not only 
amongst the most wanted and most paid jobs but also tend to be the happiest participants.  
Every student either wants to be happy or to be rich, making Cloud an interesting topic for their career. 
So why there are no courses? There are many reasons, but the most prominent is the complexity of the topic: 
To build a real cloud native application, the developer must understand virtualization, system-administration, development, 
infrastructure  and  networking. 
Additionally, the developer needs to use specific tools, know best practices, tests and work in a team.
All these factors make a high stake for entrance, also  represented in the developersurvey [22] which states that most DevOps engineers have a decade of experience, usually in operations topics.
Lucky for us, a new open-source software, Eclipse Che, is on the rise to enable everyone for cloud-native development.
Che is unlike a normal eclipse distribution: 
It is hosted on a server and developers get access with their browser. 
Instead of installing dependencies, compilers etc. locally, a common workspace is set up, which is shared amongst the programmers. 
The software build is run inside a docker-container, making it cloud-native by default. 
With Version 7 upcoming and prominent support from RedHat, Eclipse Che is not a prototype anymore. 
The latest distributions of RedHats OpenShift are shipped with Che in default. 
Therefore, it is worth looking at this possible game-changer. 
With the first part of this paper, the structure and ideas of Che are explained in more detail, the second part of this pa-
per covers general arguments about this approach and gives a detailed overview of reachable benefits for education.    
\section{Demonstrationen}
\label{sec:demos}
Das Symbol für Potenzmengen ($\powerset$) wird korrekt angezeigt.
Es ist kein Weierstraß-p ($\wp$) mehr.
